---
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
colorlinks: yes
---

# Getting started with R and Rstudio {#intro_r}

## R and Rstudio Installation

R is a language and environment for statistical computing and graphics (https://cran.r-project.org/manuals.html). Many users of R like a tool called **RStudio** (https://www.rstudio.com/). This software is what is called an Integrated Development Environment (IDE) for R. It has several nice features, including docked windows for your console and syntax-highlighting editor that supports direct code execution, as well as tools for plotting and workspace management.

### Windows operating system

- install R, https://cran.r-project.org/bin/windows/base/
- install Rstudio, https://www.rstudio.com/products/rstudio/download/#download
- [YouTube Instruction](https://www.youtube.com/watch?v=TFGYlKvQEQ4)

### macOS operating system

- install R, https://cran.r-project.org/bin/macosx/
- install Rstudio, https://www.rstudio.com/products/rstudio/download/#download (select macOS 10.14+ option)
- [YouTube Instruction](https://www.youtube.com/watch?v=LanBozXJjOk)



## R Packages
Packages are the fundamental units of reproducible R code. They include reusable R functions, the documentation that describes how to use them, and sample data.[@wickham2015r]

To load the functions in a given package, we first have to install the package. We do this using the install.packages() function. Run the line of code that installs the tidyverse package below by removing the # at the start of the second line to ‘uncomment’ the code. R will install the package to a default directory on your computer. If any dialogue box prompts you to ‘set up a personal library instead’, click yes. Once we have the package installed, we must load the functions from this library so we can use them within R.

```{r echo=T, warning=F, message=F}
  # install.packages(“tidyverse”, dependencies = T) #uncomment this line if you haven't installed this package;

  library(tidyverse) # load package library
```

The tidyverse is an opinionated collection of R packages designed for data science. All packages share an underlying design philosophy, grammar, and data structures (https://www.tidyverse.org/). The core packages are ggplot2 (data visualization), dplyr(dataframe manipulation), tidyr(data reshaping), readr(reading datasets), purrr (function and iterations) and tibble(dataframe).


### Bayesian Analysis in R using brms package

The course will mainly use the brms package in R[@burkner2017advanced], which offers a standard R-modelling type interface to the underlying computing engine Stan. Direct use of Stan is not ideal for teaching Bayesian methods.  The brms package automatically writes Stan code that can be viewed and edited, so after learning brms, the enterprising student may want to use this Stan code as a steppingstone toward programming directly in Stan. The brms package can be installed and loaded in the same way as any other R package, in this case by typing the following commands in R:

```{r echo=T, warning=F, message=F}
  # install.packages(“brms”) #uncomment this line if you haven't installed this package;
  library(brms)
```

