# Bayesian Regression I {#BayesReg1}

::: {.chapterintro data-latex=""}

- Learn how to fit linear regression model with brms
- Understand how to interpret and report regression posterior results
- Learn how to fit clustered data with brms
- Learn how to conduct model diagnostics, model comparison, and variable selection

:::

<script src="hideOutput.js"></script>

```{r, echo = FALSE, message=FALSE, warning=FALSE}
# Every chapter, we will load all the library we will use at the beginning
# of the chapter. 
library(tidyverse)
library(ggpubr)
library(brms)
library(ggmcmc)
library(MCMCpack)
library(bayesplot)
library(gtsummary)
library(coda)
library(extraDistr)
library(DT)
options(scipen = 999)
```

</br>

**What is regression?** [@mcelreath2018statistical]

- using one or more predictor variables to model the distribution of one or more outcome variables

- For a continuous outcome, we typically fit a linear regression model. 
    - Of course the relationship between outcome and predictors can be non-linear, in this case, we would consider fitting polynomial regression models or splines.
    
- For a categorical outcome, we will fit a generalized linear regression. We will cover this topic in future sessions.

- For a repeatedly measured outcome, we can fit a linear mixed-effect model (continuous outcome) or a generalized linear mixed-effect model (categorical outcome).


## Normal Models and Linear Regression

::: {.important data-latex=""}
**Conditional normal model**

- Given its mean and variance, an observation has a normal distribution

\[ Y_i \mid \mu_i, \sigma^2_i \sim N( \mu_i, \sigma^2_i) \]

- This is equivalent to the following statements

\[ Y_i = \mu_i + e_i , \ e_i \sim N(0, \sigma^2)\]

- We do not assume the collection of $Y_i, i=1, \ldots, n$ have a normal distribution
- Instead we assume the error term is normally distributed - a lesser assumption!

- In case of multiple predictors, $\mu_i$ becomes a weighted average of the $\beta_j$ values, the regression coefficient with $x_{ij}$ denoting the predictors. For example, for two covariates we have

\[ E(y_i) = \mu_i = \beta_0 + \beta_1 x_{i1} + \beta_2 x_{i2} \]

- A polynomial regression model of 2 degrees on $x_{i1}$

\[ \mu_i = \beta_0 + \beta_1 x_{i1} +  \beta_2 x_{i2} + \beta_3 x_{i1}^2 \]

- Assumptions of linear regression models
    1. **Independent observations** 
    2. **Linear relationship** We can check this assumption by examining marginal plots comparing the model predicted relationship between outcome and each continuous predictor and by examining the residual plot.
    3. **Normality of the residuals**
    4. **Homoscedasticity** Homoscedasticity in a model means that the residual is constant along the values of the dependent variable.
    5. **Multidisciplinary** Multicollinearity is the phenomenon when a number of the explanatory variables are strongly correlated.
    6. **Correctly specified regression model** This means that all relevant predictors for the response variable have been included in the model. This is often difficult to verify, but we can use posterior predictive distribution to check for regression fit.
    
:::

### One-sample and two-sample normal regression

**1. One-sample**

- If all observations have the same mean and variance, the likelihood is

\[ Y_i \mid \mu, \sigma^2 \sim N(\mu, \sigma^2)\]

- Good candidate priors for $mu$ and $\sigma$ are
    - $\mu \sim N(\text{sample mean}, \text{sample sd})$ or Student-T distributions (to handle heavy tails)
    - $\sigma \sim \text{half-t}(mean = 0, sd = \text{sample sd}, df = 3)$

**Modelling height - Textbook example on the Howell data** 

Suppose we want to fit a one-sample model on height using a dataset of anthropological measurements on a foraging people taken in the 1960s.

- First let's examine the summary statistics of variable height

<div class="fold s">
```{r echo=T, fig.align='center',fig.height=9, warning=FALSE, message=FALSE}

dat <- read.table("data/Howell.txt",header=T, sep=";")

dat %>% tbl_summary(
    statistic = list(all_continuous() ~ "{mean} ({sd})",
                     all_categorical() ~ "{n} / {N} ({p}%)"),
    digits = all_continuous() ~ 2)

dat %>% pivot_longer(everything()) %>% 
  mutate(name = factor(name, levels = c("height", "weight", "age", "male"))) %>% ggplot(aes(x = value)) +
  geom_histogram(bins = 10,fill="#008BB0",colour="#002A5C") +
  facet_wrap(~name, scales = "free", ncol = 1) +
  theme_bw()

```
<div>

- We specify our model with the following priors 
\[ heights_i \sim N (\mu, \sigma)\]
\[\mu \sim N(138, 28)\]
\[\sigma \sim \text{half-t}(0, 28, df = 3)\]

```{r echo=F, fig.cap="Prior distributions", fig.align='center', message=FALSE, warning=FALSE, fig.height=3, fig.width=5}

p1 <- tibble(x = seq(from = 60, to = 220, by = .1)) %>% 
  ggplot(aes(x = x, y = dnorm(x, mean = 138, sd = 28))) +
  geom_line() +
  scale_x_continuous(breaks = seq(from = 60, to = 220, by = 40)) +
  labs(title = "mu ~ dnorm(138, 28)",
       y = "density",
       x = "height")

p2 <- tibble(x = seq(from = 0, to = 100, by = .1)) %>% 
  ggplot(aes(x = x, y = dht(x, nu=3, sigma = 28))) +
  geom_line() +
  scale_x_continuous(breaks = seq(from = 0, to = 100, by = 10)) +
  labs(title = "sigma ~ dht(3, 28)",
       y = "density",
       x = "sigma")

ggarrange(p1,p2, nrow = 1)
```

- we now fit this model in brms

```{r echo=T, message=FALSE, warning=FALSE, eval=FALSE}
fit0 <- brm(data = dat,
             family = gaussian,
             height ~ 1, 
             prior = c(prior(normal(138, 28), class = Intercept),
                       prior(student_t(3, 0, 28), class = sigma)),
             iter = 10000, 
             warmup = 5000, 
             chains = 4, 
             cores = 4, #instructions on MCMC
             seed = 123, # random number seed to make results reproducible
             silent = 2,
             refresh = 0)

# saveRDS(fit0, "data/chap7_onesample_example")
```

- Summarizing posterior results

```{r echo=FALSE, message=FALSE, warning=FALSE}
fit0 <- readRDS("data/chap7_onesample_example")
plot(fit0)
print(fit0)
```

- Geweke index and R-hat [@geweke1991evaluating] to check for MCMC convergence

```{r echo=TRUE, message=FALSE, warning=FALSE}
# modelposterior <- as.mcmc(fit0)
# geweke.plot(modelposterior)
modelposterior <- posterior_samples(fit0)
geweke.diag(modelposterior)
```

- Posterior predictive graphic check for model fit
    - we can see here there is some skewness not captured in the model, this model is not a good fit to the data.
    - we will revisit the height model in the linear regression section where we add covariates to the model to improve the fit.

```{r echo=0, message=FALSE, warning=FALSE, fig.align='center', fig.height=3, fig.width=4.5}
pp_check(fit0, nsamples = 100)
```

::: {.guidedexercise data-latex=""}
**What is the posterior evidence that mean height is greater than 140cm?** 
```{r echo=TRUE, eval=FALSE}
hypothesis(fit0, 'Intercept > 140' , alpha = 0.025)

# proportion of posterior samples on the mean (intercept) 
# with value > 140, combining 4 chains;
modelposterior2<- posterior_samples(fit0)
sum(modelposterior2[,1]>140)/length(modelposterior2[,1])
```
:::


**2. Two-sample**

-  Simple extension of the one-sample problem
-  Observations within the each group have the same mean and variance, yielding the following likelihood
\[ Y_{ji} \mid \mu_j, \sigma^2_j \sim N(\mu_j, \sigma^2_j)\]
-  Similar candidate prior as in one-sample model
    - using normal or student-t dist and observed summary statistics
    - within-group variance should be smaller than the variance when we ignore groups
    
- We can assume equal variance $\sigma_1^2 = \sigma_2^2$ or allow them to be estimated seperately
- These two approaches are analogous to the equal-variance and unequal-variance t-tests
- In brms, this is easy to specify

**Modelling hours of sleep comparing between two drugs**

- Data which show the effect of two medication (within-patient increase in
hours of sleep) on groups consisting of 10 patients each.
- Source: [Student (1908)](https://www.jstor.org/stable/2331554?seq=1#metadata_info_tab_contents) The probable error of a mean. [@student1908probable]

- First examine the summary statistics by group

```{r echo=TRUE}
sleep <- data.frame(extra = c(0.7, -1.6, -0.2, -1.2, -0.1, 3.4, 3.7, 0.8, 0, 2, 1.9, 0.8, 1.1, 0.1, -0.1, 4.4, 5.5, 1.6, 4.6, 3.4), 
                group = c(1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 2, 2, 2, 2, 2, 2, 2, 2, 2, 2))

sleep %>% 
  mutate(group = recode_factor(group,  '1'="Drug A", '2'="Drug B")) %>%
  tbl_summary(
    by = group,
    statistic = all_continuous() ~ "{mean} ({sd})",
    digits = all_continuous() ~ 2) %>%
  add_overall()
```

- Fitting two-sample model in brms using default brms prior
    - student_t(3, 1, 2.5) for mean (same for each group)
    - student_t(3, 0, 2.5) for sigma (same for each group for unequal variance model)

```{r echo=T, message=FALSE, warning=FALSE, eval=FALSE}

fit_eq <- brm(data = sleep,
             family = gaussian,
             extra ~ group, #mean depends on group and one sigma
             prior = c(prior(normal(0, 10), class = Intercept),
                       prior(normal(0,10), class = b),
                       prior(student_t(3, 0, 2), class = sigma)),
             iter = 10000, 
             warmup = 5000, 
             chains = 4, 
             cores = 4, #instructions on MCMC
             seed = 123, # random number seed to make results reproducible
             silent = 2,
             refresh = 0)

prior_summary(fit_eq)

fit_uneq <- brm(data = sleep,
             family = gaussian,
             bf(extra ~ group, #mean depends on group
             sigma ~ group), #sigma differ by group
             prior = c(prior(normal(0, 10), class = Intercept), #prior for intercept of the extra ~ group model;
                       prior(normal(0, 10), class = b), #prior for coefficients of the extra ~ group model;
                       prior(student_t(3, 0, 2), class = Intercept, dpar = "sigma"), #prior for intercept of the sigma ~ group model;
                       prior(student_t(3, 0, 2), class = b, dpar = "sigma")), #prior for coefficients of the sigma ~ group model;
             iter = 10000, 
             warmup = 5000, 
             chains = 4, 
             cores = 4, #instructions on MCMC
             seed = 123, # random number seed to make results reproducible
             silent = 2,
             refresh = 0)

prior_summary(fit_uneq)

# saveRDS(fit_eq, "data/chap7_twosample_eq")
# saveRDS(fit_uneq, "data/chap7_twosample_uneq")
```

- Summarizing posterior results from equal variance model

```{r echo=FALSE, message=FALSE, warning=FALSE}
fit_eq <- readRDS("data/chap7_twosample_eq")
summary(fit_eq)
```

- Question: What is the probability that there is more increase in sleep with
drug A? (under equal-variance model)

```{r echo=TRUE, warning=FALSE, message=FALSE}
samples <- posterior_samples(fit_eq)
difference <- samples[,"b_group2"]
#summarize the difference
c(Prob = mean(difference > 0), Mean = mean(difference), lowCrI = quantile(difference, 0.025), highCrI = quantile(difference, 1-0.025))

#we can also use hypothesis function to evaluate this
hypothesis(fit_eq, "group2 > 0", alpha = 0.025)
```

- Summarizing posterior results from unequal variance model and evaluate the same hypothesis

```{r echo=TRUE, message=FALSE, warning=FALSE}
fit_uneq <- readRDS("data/chap7_twosample_uneq")
summary(fit_uneq)
hypothesis(fit_uneq, "group2 > 0", alpha = 0.025)
```

**How to determine equal variance or not?**

-  Approach 1, visually examine the conditional mean by group.
-  Approach 2, comparing models using information criteria (WAIC) or leave-one-out cross validation (LOO) [@vehtari2017practical]
-  **Both approaches confirm the equal variance model is the best model!**

```{r, echo = T, fig.align='center', fig.height=3, fig.width=3, message=FALSE, warning=FALSE}
plot(conditional_effects(fit_uneq), points = TRUE)

waic1 <- waic(fit_eq)
waic2 <- waic(fit_uneq)
compare_ic(waic1, waic2) #equal var best model;

loo1 <- loo(fit_eq)
loo2 <- loo(fit_uneq)
loo_compare(loo1, loo2) #equal var best model;
```

### Linear regression

-  The two-sample model is a special case of regression

\[\mu_i = \beta_0 + \beta_1 x_i\]

-  Where $x_i = 0$ for group 1 and $x_i = 1$ for group 2
    - For group 1, $\mu_i = \beta_0$, the intercept
    - For group 2, $\mu_i = \beta_0 + \beta_1$

-  brms takes care of creating dummy variables!
-  but be aware of what the reference group is and what the coefficients mean
-  levels will be by default in alphabetical order - use the factor() statement to make a variable with the desired order

::: {.guidedexercise data-latex=""}

**Try fitting the two-sample sleep data without the intercept** 

```{r echo=TRUE, eval=FALSE}
fit_eq2 <- brm(data = sleep,
             family = gaussian,
             extra ~ -1 + group, #mean depends on group and one sigma
             iter = 10000, 
             warmup = 5000, 
             chains = 4, 
             cores = 4, #instructions on MCMC
             seed = 123, # random number seed to make results reproducible
             silent = 2,
             refresh = 0)
summary(fit_eq2)
```
:::

-  In general, for a regression model

\[ y_i \mid \mu_i, \sigma^2_i \sim N(\mu_i, \sigma^2_i)\]
\[ \mu_i = \beta_0 + \beta_1 x_{i1} + \ldots + \beta_p x_{ip}\]
    -  where the $x_{ij}$ can be continuous variables or "dummy" variables
    -  the usual assumption on the variance is $\sigma_i^2 = \sigma^2$ for all subjects (Homoscedasticity - equal variance assumption)
    -  **We can seen that is technically easier to relax this assumption in a Bayesian model!** (two-sample unequal variance example)

-  Predictions
    -  We can get the predictions fro the mean value at a given value of the predictors. For example, in simple linear regression model we have
    
    \[E(Y\mid X=x) = \beta_0 + \beta_1 x\]
    - but recall that both $\beta_0$ and $\beta_1$ have posterior distributions
    - For each posterior draws of $(\beta_0, \beta_1)^s, s= 1, \ldots, S$, we have a different predicted mean
    
    \[E(Y\mid X=x, S=s) = \beta_0^s + \beta_1^s x\]
    - The set of values of these predictions across all sampled values give the posterior for the predicted mean
    
**Modelling height - Howell data revisit** 

- Let's descriptively examine the height distribution by weight, age, and sex
    - We can see that the relationship between age and height is not linear
    - we consider the following model
    
    \[heights_i \sim N(\mu_i, \sigma)\]
    \[\mu_i = \beta_0 + \beta_1 \ weight_i + \beta_2 \ age_i + \beta_3 \ age_i^2 +\beta_4 \ \text{I(sex=male)}\]

```{r echo=FALSE, fig.align='center',fig.height=3, warning=FALSE, message=FALSE}
p1 <- ggplot(aes(x = weight, y = height), data = dat) + 
  geom_point(size = 0.7) +  # add a layer with the points
  geom_smooth()

p2 <- ggplot(aes(x = age, y = height), data = dat) + 
  geom_point(size = 0.7) +  # add a layer with the points
  geom_smooth()

p3 <- ggplot(aes(x = factor(male), y = height), data = dat) + 
  geom_boxplot()  # add a layer with a boxplot

ggarrange(p1, p2, p3, nrow = 1)

```

- we consider the following prior distributions (diffuse normal on regression parameter and exponential distribution for $\sigma$)
    \[\beta_0 \sim N(0, 100) \]
    \[\beta_1 \sim N(0, 10) \]
    \[\beta_2 \sim N(0, 10) \]
    \[\beta_3 \sim N(0, 10) \]
    \[\beta_4 \sim N(0, 10) \]
    \[ \sigma \sim Gamma(1, 0.01) \]
    
```{r echo=F, fig.cap="Prior distributions", fig.align='center', message=FALSE, warning=FALSE, fig.height=3, fig.width=8}

p1 <- tibble(x = seq(from = -350, to = 350, by = .2)) %>% 
  ggplot(aes(x = x, y = dnorm(x, mean = 0, sd = 100))) +
  geom_line() +
  scale_x_continuous(breaks = seq(from = -300, to = 300, by = 100)) +
  labs(title = "N(0, 100)",
       y = "density",
       x = "Intercept")

p3 <- tibble(x = seq(from = -35, to = 35, by = .1)) %>% 
  ggplot(aes(x = x, y = dnorm(x, mean = 0, sd = 10))) +
  geom_line() +
  scale_x_continuous(breaks = seq(from = -35, to = 35, by = 5)) +
  labs(title = "N(0, 10)",
       y = "density",
       x = expression(beta))

p2 <- tibble(x = seq(from = 0, to = 1000, by = 0.2)) %>% 
  ggplot(aes(x = x, y = dgamma(x, shape = 1, rate = 0.01))) +
  geom_line() +
  scale_x_continuous(breaks = seq(from = 0, to = 1000, by = 200)) +
  labs(title = "Gamma(1, 0.01)",
       y = "density",
       x = "sigma")

ggarrange(p1,p2,p3, nrow = 1)

```
    

- we now fit this model in brms

```{r echo=T, message=FALSE, warning=FALSE, eval=FALSE}
dat <- dat %>% 
  mutate(male = factor(male, labels = c("female", "male")))

fit1 <- brm(data = dat,
             family = gaussian,
             height ~ weight + age + I(age^2) + male,
             prior = c(prior(normal(0, 100), class = Intercept),
                       # set for all "b" coefficients
                       prior(normal(0, 10), class = b),
                       prior(gamma(1, 0.01), class = sigma)),
             iter = 20000,
             warmup = 18000,
             chains = 4,
             cores = 4, #instructions on MCMC
             seed = 123, # random number seed to make results reproducible
             silent = 2,
             refresh = 0)

# saveRDS(fit1, "data/chap7_reg_example")
```

- Summarizing posterior results

```{r echo=FALSE, message=FALSE, warning=FALSE}
fit1 <- readRDS("data/chap7_reg_example")
print(fit1)
```

- Using the posterior mean, we have the following regression line

\[\hat{height}_i =  75.53 + 1.26 \ weight_i + 1.06 \ age_i -0.01 \ age_i^2 + 1.86 \ \text{I(sex=male)}\]

- Comparing between male and female of the same weight and age, the expected difference on height is estimated at 1.86cm with 95% CI[0.33, 3.41].

- What is the posterior evidence that there’s a positive association between height and sex?

```{r}
hypothesis(fit1, 'malemale > 0' , alpha = 0.025)
```


- we can visualize conditional effect of each continuous covariates on the response variable 
    - one might consider fitting $weight^2$!

```{r echo=TRUE, fig.height=3, fig.width=3}
plot(conditional_effects(fit1, "weight"),points = TRUE, point_args = list(size = 0.5))
plot(conditional_effects(fit1, "age"),points = TRUE, point_args = list(size = 0.5))
plot(conditional_effects(fit1, "male"))
```


- Coefficient of determination $R^2$ is the proportion of the variation in the dependent variable that is predictable from the independent variable(s).
    - The estimated $R^2$ is about 0.9. Really good! About 90% of the variance on height is explained by our model.

```{r echo=T}
bayes_R2(fit1)
```


- predictive intervals
    - predicting height for a female, 33 years of age and weighting 60kg

```{r echo = T}
ind_predic <- posterior_predict(fit1, newdata = data.frame(weight = 60, age = 33, male = "female"))
# mean of the predicted heights
mean(ind_predic)
# Construct a 95% posterior credible interval
posterior_interval(ind_predic, prob = 0.95)
# Posterior predictive probability of height > 164
sum(ind_predic>164)/length(ind_predic)
```


### Model diagnostics

- Geweke index and R-hat [@geweke1991evaluating] to check for MCMC convergence

```{r echo=TRUE, message=FALSE, warning=FALSE}
modelposterior<- as_draws_matrix(fit1)
geweke.diag(modelposterior)
```

- Posterior predictive graphic check for model fit
- A "good" Bayesian model produces posterior predicted sets of response values with features similar to the original data.
    - It appears our model can be improved! One suggestion is to fit a non-linear term for weight variable.
    - sometime, people also consider data transformation, e.g., log(height)
    - we can also try using student-t distribution not normal distribution to fit the mean!
    - the observed min and max height is not covered in our posterior predictive samples...
 

```{r echo=T, message=FALSE, warning=FALSE, fig.align='center', fig.height=3, fig.width=4.5}
pp_check(fit1, ndraws = 50)
pp_check(fit1, type = "stat_2d", stat = c("max", "min"))
```

- Residual plot by continuous variables - checking for linearity assumption
    
```{r echo=T, message=FALSE, warning=FALSE, fig.align='center',fig.height=3, fig.width=3}
res_df <- fit1$data %>% 
  mutate(predict_y = predict(fit1)[ , "Estimate"], 
         std_resid = residuals(fit1, type = "pearson")[ , "Estimate"])
ggplot(res_df, 
       aes(predict_y, std_resid)) + 
  labs(y="Standardized residuals") +
  geom_point(size = 0.8) + stat_smooth(se = FALSE) + theme_bw()
```

-  multicollinearity
    - If some coefficients are particularly strongly correlated, you may need to think about using a stronger prior or combining some predictors.

```{r echo=T, message=FALSE, warning=FALSE, fig.align='center'}
pairs(fit1, pars = "b", 
      off_diag_args =  # arguments of the scatterplots
        list(size = 0.5,  # point size
             alpha = 0.25))  # transparency
```


### Model comparision

::: {.important data-latex=""}

**Watanabe-Akaike Information Criteria (WAIC)**

-  WAIC [@watanabe2010asymptotic]  is an alternative approach to DIC in estimating the expected log point-wise predictive density
-  The smaller the better!
-  WAIC incorporates prior information, and the use of point-wise likelihood makes it more robust when the posterior distributions deviate from normality.
-  In general, WAIC is a better estimate of the out-of-sample deviance than AIC and DIC.
-  In R under brms package we can use `waic()` function to obtain WAIC.

**Bayesian Leave-one-out cross-validation**

-  The idea of cross-validation is to split the sample so that it imitates the scenario of estimating the parameters in part of the data and predicting the remaining part.
    -  data that are used for estimation is called the training set, and data that are used for prediction is called the validation set.

-  Leave-one-out information criteria (LOO-IC) based on posterior predictive distribution
    -  means that one uses $n-1$ observations as the training set and 1 observation as the validation sample, repeat the process $n$ times so that each time a different observation is being predicted

-  LOO-IC requires fitting the model $n$ times, it is generally very computational intensive

- WAIC can be treated as a fast approximation of LOO-IC, although LOO-IC is more robust and will be a better estimate of out-of-sample deviance

-  In R under brms package we can use `loo()` function to obtain LOO-IC.

:::

::: {.guidedexercise data-latex=""}

**Comparing model with $weight^2$ ** 

- fitting the new model on height, adjusting for age, sex, weight, $age^2$ and $weight^2$
- comparing this model to the regression model adjusted for age, $age^2$,  sex, and weight using WAIC and LOO.

```{r echo=TRUE, eval=FALSE}
fit2 <- brm(data = dat,
             family = gaussian,
             height ~ weight + I(weight^2) + age + I(age^2) + male, 
             prior = c(prior(normal(0, 100), class = Intercept),
                       prior(normal(0, 10), class = b),
                       prior(gamma(1, 0.01), class = sigma)),
             iter = 20000, 
             warmup = 18000, 
             chains = 4, 
             cores = 4, #instructions on MCMC
             seed = 123, # random number seed to make results reproducible
             silent = 2,
             refresh = 0)
saveRDS(fit2, "data/chp7_height2")

waic1 <- waic(fit1)
waic2 <- waic(fit2)
compare_ic(waic1, waic2) 

loo1 <- loo(fit1)
loo2 <- loo(fit2)
loo_compare(loo1, loo2)
```

```{r echo=TRUE, fig.height=3, fig.width=4.5}
fit2<-readRDS("data/chp7_height2")
pp_check(fit2, ndraws = 50)
```

:::


## Hierarchical models

-  Often models have more complex structures than the ones described by likelihood
and priors and simple relationships between data and parameters

-  In these models data and parameters participate in relationships of hierarchical
fashion
    -  e.g., Correlated and clustering data structures, such as measures within patients, patients within clinics etc.
    
-  Bayesian models deal with hierarchical models in a natural and intuitive way [@burkner2017advanced]

-  Source of correlation in clustered (or longitudinal) data
    - between-individual heterogeneity on i) mean and ii) velocity (rate of change; slop)
    - Within-individual/cluster homogeneity  

::: {.important data-latex=""}
**Bayesian Hierarchical Models**

-  Often Bayesian models can we written in the following layers of
the hierarchy
    1. **Data Layer** $Y \mid \theta$, the likelihood for the observed data $Y$ given model parameters $\theta$.
    2. **Adaptive/Structural Priors** $\theta \mid \alpha, \beta$, the model for parameter $\theta$ that defined by the latent data generating process governed by the hyperparameters $\alpha$ and $\beta$
    3. **Hyper Priors** $\alpha$ and $\beta$ as hyper parameters.
    
:::

Let's look at a hospital mortality study as an example. This data capture hospital mortality for a specific condition at 20 different hospitals. There is also a variable that describes the average severity of patients at each hospital.

-  First let's examine the data

```{r echo=T, message=FALSE, warning=FALSE}
dat <- read.table("data/HospitalMortality.csv",sep=",", header=T)

dat <- dat %>% 
  mutate(Hospital = factor(Hospital),
         Prop.Dead = round(Deaths/N,3))

dat %>% datatable(
  rownames = FALSE,
  options = list(
    columnDefs = list(list(className = 'dt-center', 
                      targets = 0:4))))

dat %>% select(Severity, Prop.Dead) %>% tbl_summary()
```

**1. First model - considering independent between hospital mortality**

-  We are modelling \[Deaths_i \sim Bin(p_i, n_i)\], where $i = 1, \ldots, 25$ and $p_i$ represent the risk of death for the $i$ hospital with a beta(1,1) prior on all $p_i$s
-  This model is saying that there is nothing to link mortality in hospital 1 to mortality in any other hospital
    -  We are essentially estimating the risk of mortality for each hospital

```{r echo=T, message=FALSE, warning=FALSE, eval=FALSE}
fit1 <- brm(data = dat,
            family= binomial,
            Deaths | trials(N) ~ 0 + Hospital,  
            iter = 10000,
            warmup = 8000,
            chains = 4,
            cores = 4, 
            seed = 123, 
            silent = 2,
            refresh = 0)
# saveRDS(fit1, "data/chap7_h_example1")
```

- Summarizing posterior results

```{r echo=T, message=FALSE, warning=FALSE}
fit1 <- readRDS("data/chap7_h_example1")
#posterior summary of risk (table);
round(exp(fixef(fit1)),3)
#posterior summary of risk (visual);
mcmc_intervals(fit1, transformations = "plogis")
```

**2. Second model - shared parameters on hospital mortality (complete pooling)**

-  Suppose that these were 25 hospitals in the same city
    -  Similar practices, populations, shared staff in some cases
-  How can we treat them as one big hospital with exactly the same mortality?
    -  Easy - specify in the model that they share the same probability $p$

-  Parameterizing using odds
    -  Odds are the basis for logistic regression, where we model log(odds)
    -  $0 \leq p \leq 1 \Rightarrow - \infty \leq \log(Odds) \leq \infty$ 
    -  Suggests an normal uninformative prior on log odds with a sufficient variance (ie sd = 10) 
    
    \[ \log(Odds) \sim N(0,\sigma^2)\]

-  Assume a common odds between all 25 hospitals, we run a pooled model such that \[Deaths_i \sim Bin(p, n_i)\]. This is a complete pooling model.

-  We are essentially modelling the average population risk of mortality pooling information from all 25 hospitals.

```{r echo=T, message=FALSE, warning=FALSE, eval=FALSE}
fit1.p <- brm(data = dat,
            family= binomial,
            Deaths | trials(N) ~ 1,  
            iter = 10000,
            warmup = 8000,
            chains = 4,
            cores = 4, 
            seed = 123, 
            silent = 2,
            refresh = 0)
# saveRDS(fit1.p, "data/chap7_h_example1p")
```

- Summarizing posterior results

```{r echo=F, message=FALSE, warning=FALSE}
fit1.p <- readRDS("data/chap7_h_example1p")
#posterior summary of risk overall;
round(exp(fixef(fit1.p)),3)
```

-  Drawbacks of a complete pooling approach

There are two main drawbacks to taking a complete pooling approach to analyze group structured data:

-  in case of clustered and correlated data, we violate the assumption of independence
-  misleading conclusions about the relationship between and across subjects

**3. third model - shared parameters on hospital mortality (partial pooling)**

-  Suppose that these were 25 hospitals in the same city
    -  Similar (but **not same**) practices and patient populations; partly shared staff
-  How can we treat them as similar
    -  with a common underlying true mortality
    -  with **random variation** from hospital to hospital in the true mortality!
    
-  Need to introduce two parameters representing
    -  Common mortality (mean log odds)
    -  Hospital-to-hospital variation (sd of log odds)
    
-  This introduces an **hierarchical structure** in the model!

-  Specifically, we have
    1. **Data Layer** $y_i \mid p_i \sim P(y_i \mid p_i)$
    2. **Adaptive/Structural Priors** $logit(p_i) \mid \theta, \sigma^2 \sim P(logit(p_i) \mid \theta, \sigma^2)$
    3. **Hyper Priors** $\theta \sim P(\theta)$ and $\sigma^2 ~ P(\sigma^2)$ as hyper parameters.

-  A Bayesian random intercept model!

```{r echo=T, message=FALSE, warning=FALSE, eval=FALSE}
fit1.ri <- brm(data = dat,
            family= binomial,
            Deaths | trials(N) ~ 1 + (1|Hospital),  
            iter = 10000,
            warmup = 8000,
            chains = 4,
            cores = 4, 
            seed = 123, 
            silent = 2,
            refresh = 0)
# saveRDS(fit1.ri, "data/chap7_h_example1ri")
```

- Summarizing posterior results on risk and between-hospital standard deviation of the intercept

```{r echo=T, message=FALSE, warning=FALSE}
fit1.ri <- readRDS("data/chap7_h_example1ri")
summary(fit1.ri)

#posterior summary of risk overall (table);
round(exp(fixef(fit1.ri)),3)

s1 <- posterior_samples(fit1.ri)
sigma <- s1[,"sd_Hospital__Intercept"]
qs <- quantile(sigma,c(0.025, 0.975))

temp <- density(sigma,n = 1024)
dp1 <- data.frame(sigma=temp$x, density=temp$y, area=cut(temp$x, c(0,qs,10), include.lowest = T))

ggplot(dp1, aes(x=sigma, y=density))+geom_area(aes(fill=area), show.legend=FALSE)+
    theme_bw()+
    geom_line(color="orange")+
    ggtitle(paste0("Mean sigma= ", round(mean(sigma),2),"; 95% CrI: ", round(qs[1],1),
                   " to ",round(qs[2],2)))+
    scale_fill_manual(values=c("orange","white","orange"))

```


-  Hospitals, ranked best to worst by posterior odds of mortality (median ranking on mortality, along with a 95% CrI)

```{r echo = T}
h <- s1[,grep(",Intercept",names(s1))]
r <- apply(h,1,rank)
rankings <- t(apply(r,1,quantile, c(0.5, c(0.025, 0.975))))
row.names(rankings) <- paste("Hospital",1:25)
colnames(rankings)[1] <- "Median"
rankings <- data.frame(rankings, ObservedProportion=round(dat$Deaths/dat$N,3), n=dat$N)
rankings <- rankings[order(rankings[,1]),]
knitr::kable(rankings, caption="Hospitals, ranked best to worst by posterior odds of mortality")
```



### Clustered data (Multi-level)

-  Example: Developmental toxicity study of ethylene glycol (EG) [@price1985developmental]

-  The data are from a development toxicity study of ethylene glycol (EG). 

-  Ethylene glycol is a high-volume industrial chemical used in many applications. 

-  In a study of laboratory mice conducted through the National Toxicology Program (NTP), EG  was administered  at doses of 0, 750, 1500, or 3000 mg/kg/day to 94 pregnant mice (dams) over the period of major organogenesis, beginning just after implantation. 

- Following sacrifice, fetal weight and evidence of malformations were recorded 

-  Let's first read in the data
    -  This data set is in a long format
    -  we can see each row represents a mouse.
    -  Mice are nested within the litter. **We have a two-level data!**
    
<div class="fold s">
```{r echo = T}
ntp <- read.table("data/ntp-data.txt", header = TRUE)
colnames(ntp) <- c("Litter", "Dose", "Weight", "Malformation")
ntp <- ntp %>%
  group_by(Litter) %>%
  mutate(Litid = row_number()) %>% # seq id for each litter
  mutate(FetalID = paste0(Litter, ".", Litid)) %>% # FetalID = LitID.tempid
  select(-Litid) # remove tempid from toxic2
knitr::kable(head(ntp), row.names = F)
```
<div>

-  Number of doses in each group and number of fetuses in each litter

```{r echo = T}
cluster <- ntp %>% group_by(Dose) %>%
  summarize(nlitter = n_distinct(Litter), nfetal = n_distinct(FetalID))
cluster %>% datatable(
  rownames = FALSE,
  options = list(
    columnDefs = list(list(className = 'dt-center', 
                      targets = 0:2))))

clustersize <- ntp %>% group_by(Litter) %>% summarise(csize = n_distinct(FetalID))
# clustersize %>% datatable(
#   rownames = FALSE,
#   options = list(
#     columnDefs = list(list(className = 'dt-center', 
#                       targets = 0:1))))

clustersize %>%
ggplot(aes(x = Litter, y = csize)) +
geom_bar(stat = "identity", fill = "black", color = "snow1")+
  theme_bw()

```

-  Visualize relationship between fetal malformation and dose
```{r message=FALSE, warning=FALSE}
ntp %>%
  group_by(Dose, Litter) %>%
  summarise(fmalprop = mean(Malformation)) %>%
  ggplot(aes(x = Litter, y = fmalprop)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  ylab("Proportion of Malformation")+
  facet_wrap(~Dose)+
  theme_bw()
```


-  Visualize relationship between fetal weight and dose
```{r message=FALSE, warning=FALSE}
ntp %>%
  ggplot(aes(x = Litter, y = Weight)) +
  geom_boxplot(aes(group = Litter), color = "steelblue") +
  facet_wrap(~Dose)+
  theme_bw()

ntp <- ntp %>% 
  mutate(Dose = factor(Dose, levels = c("0","750","1500","3000")))
```

-  Suppose we are interested to assess the effect of dose on fetal weight.
-  The data on fetal weight from this experiment are clustered, with observations on the fetuses nested within litters.

-  **First model - Normal hierarchical model with random intercept**

-  data structure: model of fetal weight within each litter (index by j)
\[ Y_{ij} \mid \mu_j, \sigma_y \sim N(\mu_j, \sigma_y^2)\]

-  adaptive priors on litter-level mean

\[\mu_j \mid \mu, \sigma_\mu \sim N(\mu, \sigma_u^2)\]

-  hyperparameter priors on $\mu$, $\sigma_u$ as well as global prior for $\sigma_y$


-  Variability within litter by dose
    -  in this model, only $\beta_{0j}$ depends on litter 
    -  $\beta_{1}$ is shared across all litter $j$, the expected change in fetal weight comparing between various dosage to reference level (dose = 0)
    -  $\sigma_y$ measures **Within-group variability** around the mean regression model $\beta_{0j} + \beta_{1} \ dose_{ij}$

\[\mu_{ij} = \beta_{0j} + \beta_{1} \ dose_{ij}\]

-  specifying our hierarchical random intercepts model (update with $\beta$):

\[ Y_{ij} \mid \beta_{0j}, \beta_{1}, \sigma_y \sim N(\beta_{0j} + \beta_{1} \ dose_{ij}, \sigma_y^2) \]

\[\beta_{0j} \mid \beta_{0}, \sigma_0^2\]

\[\beta_{0} \sim N(0,10)\]
\[\beta_{1} \sim N(0,10)\]
\[\sigma_y \sim HalfCauchy(10)\]
\[\sigma_0 \sim HalfCauchy(10)\]


```{r echo=TRUE, eval=FALSE}
fit2 <- brm(data = ntp,
            family= gaussian(),
            Weight ~ Dose + (1|Litter),  
            prior = c(prior(normal(0, 10), class = Intercept),
                      prior(normal(0, 10), class = b),
                      prior(cauchy(0, 10), class = sd),
                      prior(cauchy(0, 10), class = sigma)),
            iter = 10000,
            warmup = 8000,
            chains = 4,
            cores = 4, 
            seed = 123, 
            silent = 2,
            refresh = 0)
# saveRDS(fit2, "data/chap7_h_example2")
```


```{r echo = T, fig.align='center', fig.height=9}
fit2 <- readRDS("data/chap7_h_example2")
summary(fit2)
# checking traceplot
# fit2 %>% plot( combo = c("hist", "trace"), widths = c(1, 1.5),
#       theme = theme_bw(base_size = 10))
conditional_effects(fit2)
```


-  The adjusted ICC represents the proportion of variance of the outcome that are due to between-level (e.g., between-group, between-person) differences
    -  **Does not consider fixed effect!** In this case, adjusted ICC does not feature variation caused by dose.

\[\rho_{adj} = \frac{\sigma_0^2}{\sigma_0^2 + \sigma^2}\]

-  where $\sigma_0^2$ is the between-level SD, which is the SD of cluster means (i.e., variability between litters) and $\sigma$ is the observation-specific SD (i.e., variability within a litter)

-  The conditional ICC considers variation with fixed effect, in this case dose.

\[\rho_{c} = \frac{\sigma_0^2}{\sigma_0^2 + \sigma^2 + \sigma_f^2}\]

\[\sigma_f^2 = var(\sum_{m=1}^p \beta_m x_{mij}), p \text{ # of coefficients exclude } \beta_0\]

-  reference: [@nakagawa2017coefficient] [doi: 10.1098/rsif.2017.0213](https://royalsocietypublishing.org/doi/10.1098/rsif.2017.0213)

-  We can use the posterior draws of the two SD to compute the posterior for the adjusted ICC:

```{r echo=TRUE, warning=FALSE, message=FALSE}
# Computing ICC
# 1. Obtain posterior draws of tau and sigma
sd_m2 <- VarCorr(fit2, summary = FALSE)
draws_tau <- sd_m2$Litter$sd[ , "Intercept"]  # tau
draws_sigma <- sd_m2$residual__$sd[ , 1]  #sigma
# 2. Compute draws for adjusted ICC
draws_icc <- draws_tau^2 / (draws_tau^2 + draws_sigma^2)
# quantile of posterior adjusted ICC
quantile(draws_icc, p=c(0.5, 0.025,0.975))   

# Plot the adjusted ICC
qplot(draws_icc, geom = "density", xlab = "ICC adjust", bw = "SJ")
```

-  It was estimated that 56%, 95% CI [48%, 65%] of the variations in fetal weight was attributed to between-litter differences.


-  We can use the posterior draws of b_Dose750, b_Dose1500, and b_Dose3000 (fix effects) to compute the posterior for the conditional ICC:

```{r echo=TRUE, warning=FALSE, message=FALSE}
# sjstats::icc(fit2) point estimate of adjusted and conditional ICC;
# Intraclass Correlation Coefficient
# 
#      Adjusted ICC: 0.565
#   Conditional ICC: 0.317

# calculating conditional ICC
post_par <- colMeans(posterior_samples(fit2, pars = c("b_Dose750","b_Dose1500","b_Dose3000")))
dummy <- with(ntp, outer(Dose, levels(Dose), `==`)*1)[,-1] #dummy matrix removing dose0 level;
sigma_f2 <- var(dummy%*%post_par)
draws_icc_conditional <- (draws_tau^2) / (draws_tau^2 + sigma_f2+ draws_sigma^2)
# quantile of posterior conditional ICC
quantile(draws_icc_conditional, p=c(0.5, 0.025,0.975)) 
# Plot the adjusted ICC
qplot(draws_icc_conditional, geom = "density", xlab = "ICC conditional", bw = "SJ")
```


-  It was estimated that 31%, 95% CI [25%, 40%] of the variations in fetal weight was attributed to between-litter differences given fix effects.


-  **Second model - Normal hierarchical model with varying intercept and slope**

-  Variability in "velocity" between individual, dose the effect of dose on weight also vary by litter?

-  We want to build a model which recognizes that in the relationship between feta weight and dose, 
    -  where the intercept (i.e., baseline weight) 
    -  and slope (i.e., rate at which weight changes with dose) might vary from mouse to mouse.
-  A random intercept and slope model!

-  Variability within litter by dose
    -  in this model, $\beta_{0j}$ and $\beta_{1j}$ depend on litter 
  
-  specifying our hierarchical random intercept and slope model :

\[ Y_{ij} \mid \beta_{0j}, \beta_{1j}, \sigma_y \sim N(\beta_{0j} + \beta_{1j} \ dose_{ij}, \sigma_y^2) \]

\[\beta_{0j} \mid \beta_{0}, \sigma_0^2\]
\[\beta_{1j} \mid \beta_{1}, \sigma_1^2\]

-  In this model $\beta_{0j}$ and $\beta_{1j}$ working together to describe the model for litter $j$, and thus are **correlated**!

-  Let $\rho \in [-1,1]$ represents the correlation between $\beta_{0j}$ and $\beta_{1j}$. We can assign a **joint Normal model** of $\beta_{0j}$ and $\beta_{1j}$ where

\[
\begin{pmatrix}
\beta_{0j}\\
\beta_{1j}
\end{pmatrix} \mid \beta_{0}, \beta_{1}, \sigma_0, \sigma_1 \sim N(\begin{pmatrix}
\beta_{0}\\
\beta_{1}
\end{pmatrix}, \begin{pmatrix}
\sigma_0^2 & \rho \sigma_0\sigma_1 \\
\rho \sigma_0\sigma_1 & \sigma_1^2 
\end{pmatrix}).
\]

\[\beta_{0} \sim N(0,10)\]
\[\beta_{1} \sim N(0,10)\]
\[\sigma_y \sim HalfCauchy(10)\]
\[\sigma_0 \sim HalfCauchy(10)\]
\[\sigma_1 \sim HalfCauchy(10)\]
\[\mathbf{R} \sim LKJcorr(2)\]

-  where $\mathbf{R}$ is the correlation matrix $$\begin{pmatrix}
1 & \rho  \\
\rho  & 1 
\end{pmatrix}$$

-  LKJ-correlation prior [@lewandowski2009generating] with one parameter ($\zeta$) that controls the strength of the correlation.

```{r echo=T, fig.cap="LKJ prior for different values of the shape paramete", fig.align='center', fig.width=9, warning=FALSE, message=FALSE}
dlkjcorr2 <- function(rho, eta = 1, log = FALSE) {
  # Function to compute the LKJ density given a correlation
  out <- (eta - 1) * log(1 - rho^2) - 
    1 / 2 * log(pi) - lgamma(eta) + lgamma(eta + 1 / 2)
  if (!log) out <- exp(out)
  out
}
ggplot(tibble(rho = c(-1, 1)), aes(x = rho)) + 
  stat_function(fun = dlkjcorr2, args = list(eta = 0.1), 
                aes(col = "0.1"), n = 501) + 
  stat_function(fun = dlkjcorr2, args = list(eta = 0.5), 
                aes(col = "0.5"), n = 501) + 
  stat_function(fun = dlkjcorr2, args = list(eta = 1), 
                aes(col = "1"), n = 501) + 
  stat_function(fun = dlkjcorr2, args = list(eta = 2), 
                aes(col = "2"), n = 501) + 
  stat_function(fun = dlkjcorr2, args = list(eta = 10), 
                aes(col = "10"), n = 501) + 
  stat_function(fun = dlkjcorr2, args = list(eta = 100), 
                aes(col = "100"), n = 501) + 
  labs(col = expression(eta), x = expression(rho), y = "Density")+
  theme_bw()
```


```{r echo=TRUE, eval=FALSE}
fit3 <- brm(data = ntp,
            family= gaussian(),
            Weight ~ Dose + (Dose|Litter),  
            prior = c(prior(normal(0, 10), class = Intercept),
                      prior(normal(0, 10), class = b),
                      prior(cauchy(0, 10), class = sd),
                      prior(cauchy(0, 10), class = sigma),
                      prior(lkj(2), class = cor)),
            iter = 11000,
            warmup = 10000,
            chains = 3,
            cores = 4, 
            control = list(adapt_delta = .9), 
            seed = 123, 
            silent = 2,
            refresh = 0)
# saveRDS(fit3, "data/chap7_h_example3")
```

-  Posterior summary

```{r echo = T, warning=FALSE, message = FALSE, fig.align='center', fig.height=9}
fit3 <- readRDS("data/chap7_h_example3")
summary(fit3)
fit3 %>% plot( combo = c("hist", "trace"), widths = c(1, 1.5),
      theme = theme_bw(base_size = 10))
conditional_effects(fit3)
```

-  Interpretations

```{r}
round(VarCorr(fit3)$Litter$sd, 3)
```



-  Comparing between the two models (both models are really similar)

```{r message=FALSE, warning=FALSE}
loo(fit2, fit3)

waic1 <- waic(fit2)
waic2 <- waic(fit3)
compare_ic(waic1, waic2) 

```


## Regression regularization

### Shrinkage Priors in action

-  When the number of parameters to be estimated is large relative to sample size, estimation using non-informative or weakly informative priors tend to overfit.

-  One way to avoid overfitting is to perform regularization, that is, to shrink some of the parameters to closer to zero. This makes the model fit less well to the existing data, but will be much more generalizable to an independent data set.

-  Sparsity-Inducing Priors

-  The horseshoe prior is a type of hierarchical prior for regression models by introducing a global scale and local scale parameters on the priors for the regression coefficients. 

-  Specifically, with $p$ predictors

\[Y_i \sim N(\mu_i, \sigma^2)\]
\[\mu_i = \beta_0 + \sum_{m=1}^p\beta_m X_m\]
\[\beta_0 \sim N(0,10)\]
\[\beta_m \sim N(0,\tau\lambda_m)\]
\[\lambda_m \sim Cauchy(10)\]
\[\tau \sim Cauchy(\tau_0)\]

-  local scale $\lambda_m$ can flexibly shrink the coefficient to close to zero.

```{r warning=FALSE,message=FALSE, fig.align='center'}
dhs <- Vectorize(
  function(y, df = 1) {
  ff <- function(lam) dnorm(y, 0, sd = lam) * dt(lam, df) * 2
  if (y != 0) integrate(ff, lower = 0, upper = Inf)$value
  else Inf
  }
)
ggplot(data.frame(x = c(-6, 6)), aes(x = x)) +
  stat_function(fun = dhs, args = list(df = 1), n = 501,
                aes(col = "HS"), linetype = 1) +
  stat_function(fun = dnorm, n = 501, 
                aes(col = "norm"), linetype = 2) +
  scale_color_manual("", values = c("steelblue", "black"),
                     labels = c("horseshoe(1)", "N(0, 1)")) +
    xlab("y") + ylab("density") + ylim(0, 0.75)+
  theme_bw()
```


### Bayesian logistic regression model (COPE example revisit)

- Data come from a cross-sectional study of 1,225 smokers over the age of 40.

- Each participant was assessed for chronic obstructive pulmonary disease (COPD), and characteristics of the type of cigarette they most frequently smoke were recorded. 

- The objective of the study was to identify associations between COPD diagnosis and cigarette characteristics

We will use the following variables from this data set:

- NIC: Nicotine content, in mg

- TAR: Tar content, in mg

- LEN: Length of cigarette, in mm

- copd: 1 = COPD diagnosis, 0 = no COPD diagnosis

```{r echo=F}
copd <- read.csv("data/copd.csv", header = T)
copd <- copd %>% select("NIC","TAR","LEN","copd")
knitr::kable(head(copd))
```

- We fit a simple logistic regression, predicting the risk of copd by all covariates as well as all interaction effects

```{r echo=T, message=FALSE, warning=FALSE, eval=FALSE}
fit4.1 <- brm(copd ~ (.)^2, 
               data = copd, 
               family = bernoulli(link = "logit"), 
               prior = c(
                 prior(normal(0,10), class = "Intercept"),
                 # Prior guess of 25% of the terms are non-zero
                 prior(horseshoe(par_ratio = 2 / 8), class = "b")),
            iter = 7500,
            warmup = 5000,
            chains = 4,
            cores = 5, 
            seed = 123, 
            silent = 2, 
            refresh = 0)

fit4.2 <- brm(copd ~ (.)^2, 
               data = copd, 
               family = bernoulli(link = "logit"), 
               prior = c(
                 prior(normal(0,10), class = "Intercept"),
                 prior(normal(0,10), class = "b")),
            iter = 7500,
            warmup = 5000,
            chains = 4,
            cores = 5, 
            seed = 123, 
            silent = 2, 
            refresh = 0)
# saveRDS(fit4.1, file="data/chap7_h_example4.1")
# saveRDS(fit4.2, file="data/chap7_h_example4.2")
```

```{r echo=T, message=FALSE, warning=FALSE}
fit4.1 <- readRDS("data/chap7_h_example4.1")
fit4.2 <- readRDS("data/chap7_h_example4.2")

summary(fit4.2)
summary(fit4.1)
```


-  An arbitrary cutoff is to select only coefficients with posterior means larger than .02, in which case only NIC and TAR and their interaction were supported by the data.

### R Session information {-}

```{r echo=F}
getS3method("print","sessionInfo")(sessionInfo()[c(1:7)])
```



